### 2021-11-22 scratch.tex

\textcolor{red}{
\begin{equation}
    \small
    \vb*{M_{ij}} = 
    \frac{\langle e^{+}_{i}/R_{ii} \rangle - \langle e^{-}_{i}/R_{ii} \rangle}{\abs{\Delta g_{j}}} - \delta_{ij} 
    = \frac{\langle e^{obs,+}_{i}\rangle - \langle e^{obs,-}_{i}\rangle}{\abs{\Delta g_{j}}} - \delta_{ij}, 
\end{equation}
where the numerator calculates the \textsc{metacalibration} calibrated observed shear by dividing $i$-th component of the ellipticity of the unsheared \textsc{metacalibration} catalog by shear response.
\begin{equation}
    \small
    \vb*{C_{ij}} = \frac{\langle e^{obs,+}_{i} - (\delta_{ij}+M_{ij})e^{true,+}_{i})\rangle + \langle e^{obs,-}_{i} - (\delta_{ij}+M_{ij})e^{true,-}_{i}) \rangle}{2}
\end{equation}